---
title: "p8105_hw2_hj2660"
author: "Hyun Jin Jung"
date: "2023-10-04"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(dplyr)
```

## Problem 1

Clean the data in `pols-month.csv`. There are months when Ford became president following Nixon, during which the `prez_gop` variable is 2. We can reassign these cases to be coded as `gop`, which is consistent with months when `prez_gop` is 1. This ensures that the `president` variable maintains consistency in labeling the political affiliation, regardless of the transition between Nixon and Ford.

```{r}
pols_df = 
  read.csv("fivethirtyeight_datasets/pols-month.csv") |>
  separate(mon, into = c("year", "month", "day"), sep = "-") |>
  mutate(month = month.abb[as.integer(month)]) |>
  mutate(president = recode(prez_gop, "0" =  "dem", "1" = "gop", "2" = "gop")) |>
  mutate(year = as.integer(year)) |>
  select(-prez_gop, -prez_dem, -day)
```

Clean the data in `snp.csv`, and to ensure the year is in four digits, we can use `if_else`.

```{r}
snp_df = 
  read.csv("fivethirtyeight_datasets/snp.csv") |>
  separate(date, into = c("month", "day", "year"), convert = TRUE) |>
  select(year, month, everything()) |>
  mutate(month = month.abb[as.integer(month)]) |>
  mutate(year = if_else(year >= 50, 1900 + year, 2000 + year)) |>
  mutate(year = as.integer(year)) |>
  select(year, month, close)
```

Tidy data in `unemployment.csv` and change from 'wide' to 'long' format so that it can be merged with the other two datasets.

```{r}
unemployment_df = 
  read.csv("fivethirtyeight_datasets/unemployment.csv") |>
  rename(year = Year) |>
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment"
  ) |>
  select(year, month, unemployment)
```

Join the datasets by merging `pols_df`, `snap_df`, and `unemployment_df`.

```{r}
data = 
  left_join(pols_df, snp_df, by = c("year", "month")) |>
  left_join(unemployment_df, by = c("year", "month"))

str(data)
```

```{r, echo=FALSE, results='hide'}
average_unemployment_dem = data |>
  filter(month == "Jan", year >= 1975, president == "dem") |>
  pull(unemployment) |>
  mean() |>
  round(2)
```

```{r, echo=FALSE, results='hide'}
average_unemployment_gop = data |>
  filter(month == "Jan", year >= 1975, president == "gop") |>
  pull(unemployment) |>
  mean() |>
  round(2)
```

The `pols_df` has `r nrow(pols_df)` observations and `r ncol(pols_df)` variables and tells about the party affiliation distribution of national politicians from 1947 to 2015. The `snp_df` has `r nrow(snp_df)` observations and `r ncol(snp_df)` variables, ranging from years 1950 to 2015. The `unemployment_df` has `r nrow(unemployment_df)` observations and `r ncol(unemployment_df)` variables, covering the years from 1948 to 2015. The average unemployment rate for January, starting in the year 1975, during months when the president is a Democrat, is `r average_unemployment_dem`%. The average unemployment rate for Republicans over the same time period was `r average_unemployment_gop`%. Missing values in the 'close' and 'unemployment' variables have been replaced with 'NA' values.


## Problem 2


Clean and organize the dataset for Mr. Trash Wheel.
```{r}
MrTrashWheel_df = 
  readxl::read_excel("./Trash_Wheel/202309 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "B2:N586") |>
  janitor::clean_names() |>
  mutate(homes_powered = (weight_tons * 500)/30) |>
  mutate(trash_wheel = "Mr. Trash Wheel") |>
  mutate(year = as.character(year)) 
```


Clean and organize the dataset for Professor Trash Wheel.
```{r}
ProfTrashWheel_df =
  readxl::read_excel("./Trash_Wheel/202309 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "B2:M108") |>
  janitor::clean_names() |>
  mutate(homes_powered = (weight_tons * 500)/30) |>
  mutate(trash_wheel = "Professor Trash Wheel") |>
  mutate(year = as.character(year))
```

Clean and organize the dataset for Gwynnda Trash Wheel.
```{r}
GwynTrashWheel_df =
  readxl::read_excel("./Trash_Wheel/202309 Trash Wheel Collection Data.xlsx", sheet = "Gwynnda Trash Wheel", range = "B2:K157") |>
  janitor::clean_names() |>
  mutate(homes_powered = (weight_tons * 500)/30) |>
  mutate(trash_wheel = "Gwynnda Trash Wheel") |>
  mutate(year = as.character(year))
```

While combining three datasets, an error occurred, indicating that there was a data type mismatch between the `year` columns, with one being of type 'character' and the other of type 'double.' To resolve this issue, the `year` column in each dataset was uniformly converted to the 'character' data type. Additionally, a `trash_wheel` column was introduced in each dataset to specify the corresponding Trash Wheel.

```{r}
combined_TW_df = 
  bind_rows(MrTrashWheel_df, ProfTrashWheel_df, GwynTrashWheel_df)
```

In the `MrTrashWheel_df` dataset, there are `r nrow(MrTrashWheel_df)` observations and `r ncol(MrTrashWheel_df)` variables. The average number of homes powered is `r MrTrashWheel_df |> pull(homes_powered) |> mean()`, and the total weight of trash collected is `r MrTrashWheel_df |> pull(weight_tons) |> sum()` tons. For the `ProfTrashWheel_df` dataset, it consists of `r nrow(ProfTrashWheel_df)` observations and `r ncol(ProfTrashWheel_df)` variables, with an average of homes powered equal to `r ProfTrashWheel_df |> pull(homes_powered) |> mean()`. The total weight of trash collected by Professor Trash Wheel is `r ProfTrashWheel_df |> pull(weight_tons) |> sum()` tons. For the `GwynTrashWheel_df` dataset, there are `r nrow(GwynTrashWheel_df)` observations and `r ncol(GwynTrashWheel_df)` variables. The average number of homes powered is `r GwynTrashWheel_df |> pull(homes_powered) |> mean()`, and the total weight of trash collected is `r GwynTrashWheel_df |> pull(weight_tons) |> sum()` tons. Lastly, Gwynnda collected a total of `r GwynTrashWheel_df |> filter(format(date, "%Y-%m") == "2021-07") |> pull(cigarette_butts) |> sum() |> as.integer()` cigarette butts in July of 2021.

To confirm the accuracy of the combined data, I counted the `Mr. Trash Wheel` entries in the `combined_TW_df` dataset, which is `r combined_TW_df |> filter(trash_wheel == "Mr. Trash Wheel") |> summarise(count = n())`. The result matches the number of records in the `MrTrashWheel_df` dataset, ensuring that the data is correct.

## Problem 3

```{r}
MCI_baseline_df = 
  read.csv("data_mci/MCI_baseline.csv", skip = 1, na = ".") |>
  janitor::clean_names() |>
  mutate(sex = recode(sex, '1' = "male", '0' = "female")) |>
  mutate(apoe4 = recode(apoe4, '1' = "APOE4 carrier", '0' = "APOE4 non-carrier"))

dev_MCI = 
  MCI_baseline_df |>
  filter(age_at_onset !=".")
```

```{r}
apoe4_carriers_w =
  MCI_baseline_df |>
  filter(sex == "female") |> 
  filter (apoe4 == "APOE4 carrier") |> 
  nrow()

total_female = 
  MCI_baseline_df |>
  filter(sex == "female") |>
  nrow()

proportion_apoe4_carriers =
  apoe4_carriers_w / total_female
```

Import `MCI_baseline.csv` and in order to read csv file from row 2, used `skip = 1` and changed '.' to 'NA'. Use `mutate` function to update the `sex` and `apoe4` column. `filter` function has been used to remove the rows where `age_at_onset` has empty values which indicates no MCI at baseline. `r nrow(MCI_baseline_df)` participants were recruited and `r nrow(dev_MCI)` developed MCI. The average baseline age of all the participants is `r MCI_baseline_df |> pull(current_age) |> mean()` years old and the average baseline age for ones who developed MCI is `r dev_MCI |> pull(current_age) |> mean()` years old. Number of women who are APOE4 carriers is `r apoe4_carriers_w` and the number of total female is `r total_female`. The proportion of women in the study that are APOE4 carriers is `r proportion_apoe4_carriers`.

```{r}
mci_amyloid_df = 
  read.csv("data_mci/mci_amyloid.csv", skip = 1, , na = "Na") |>
  janitor::clean_names() 

mci_amyloid_long =
  pivot_longer(mci_amyloid_df,
               cols = baseline:time_8,
               names_to = 'time',
              values_to = 'value') |>
  mutate(time = recode(time,
    'baseline' = 0,
    'time_2' = 2,
    'time_4' = 4,
    'time_6' = 6,
    'time_8' = 8
  )) |>
  rename(id = study_id)
```

Following the same procedure as with `mci_amyloid_df.csv`, skip the first row (header) using `skip = 1`. There were 'Na' value and 'NA', so I converted using `na = "Na"` for consistency. Since this dataset contains longitudinal data, pivot it from wide format to long format using `pivot_longer` for improved analysis. To enhance dataset clarity, rename the `time` column with more descriptive labels using `recode`. To combine the demographic and biomarker datasets, rename `study_id` to `id`.

Combine demographic and biomarker datasets and export the result as a CSV!!
```{r}
combined_MCI_df = 
  dev_MCI |>
  inner_join(mci_amyloid_long, by = "id")

write_csv(combined_MCI_df, "data_mci/combined_MCI_df.csv")
```

The `id` range for `MCI_baseline_df` dataset is [`r range(MCI_baseline_df$id)`] and has `r nrow(MCI_baseline_df)` observations. The `study_id` range for `mci_amyloid_df` dataset is [`r range(mci_amyloid_df$study_id)`] and has `r nrow(mci_amyloid_df)` observations. So, `id` 484 to 495 is missing from `MCI_baseline_df` dataset and some of the data ranging from 1 to 483 is missing from `mci_amyloid_df`.

In order to combine the demographic and biomarker datasets and retain only the participants who appear in both datasets, we used `inner_join` to merge `dev_MCI` based on the `id` with `mci_amyloid_long`, which we pivoted to long format. Combining the original data `MCI_baseline_df` seemed meaningless since it contained those who did not develop MCI.

The average Aβ 42/40 ratio (`value` column) for participants classified as `APOE4 carrier` in the dataset is `r combined_MCI_df |> filter(apoe4 == "APOE4 carrier") |> summarise(mean_value = mean(as.numeric(value), na.rm = TRUE))` and the mean value for those who are `APOE4 non-carrier` is `r combined_MCI_df |> filter(apoe4 == "APOE4 non-carrier") |> summarise(mean_value = mean(as.numeric(value), na.rm = TRUE))`. `APOE4 carrier` tend to have altered amyloid metabolism, resulting in lower Aβ 42/40 ratios, which explains the increased risk of developing Alzheimer's disease in APOE4 carriers.



